---
title: "Evaluating length-weight of pallid sturgeon."
author: Michael E. Colvin
date: 4 October 2016
output:
  md_document:
    variant: markdown_github
  html_document:
    theme: flatly
    highlight: espresso
    toc: FALSE
---



```{r,echo=FALSE}
## Analysis preliminaries
library(pander)
#setwd(file.path(Sys.getenv("USERPROFILE"),"My Documents/projects/pallid sturgeon/analysis/skinny-fish"))
source("./R/1_global.R")
source("./R/2_functions.R")
#source("./R/3_load.R")
#source("./R/4_clean.R")
dat<-readRDS("dat/dat.RDS")
source("./R/5_figures.R")
source("./R/6_tables.R")
source("./R/7_analysis.R")


```
## Analysis source

Analysis source code can be viewed at [https://github.com/mcolvin/PDSG-LW](https://github.com/mcolvin/PDSG-LW).



## Length weight analysis of covariance

### Analysis overview:

This analysis fits length-weight data for pallid sturgeon captured in the upper (RPMA 2 and 3) and
lower (RPMA 4) basins of the Missouri River.  The rationale for this analysis is that condition is 
an index of weight given an expected weight and therefore modelling expected weight as a function
of length using the base model $log(Weight) = log(a)+b\cdot log(Length)$ will illustrate changes in
expected weight. This approach allows evaluating how expected weight is changing over time that
complements analyses of condition. Specifically, if conditions is decreasing over time, it should index
a decrease in expected fish weight for a given length. 

### Methods in a nutshell

1. Fit length-weight data as a linear model by log transforming length and weight
1. Estimate effect of year, basin and specific interactions
1. Year is treated as a factor rather than continuous.
1. Use AICc model averaging to evaluate evidence for differences among year and basins
1. Used 'top' model to predict weights and 95% confidence intervals for 800 (i.e., reproductive fish)
to 1400 mm pallid sturgeon by increments of 200. 



### Question 1. Does the length-weight relationship vary among years and between basins?

Short answer, yes.  There is overwhelming evidence that length weight model parameters ($a$ and $b$)
should vary among year and basin.  The AICc model selection table below provides the support.

```{r, echo=FALSE}
pander(tables(2),rownames=FALSE)
```

The model selection results indicate that the length weight relationship should be viewed like
the figures 1 and 2 below where there is essentially a separate length-weight relationship for each year
and basin.  

```{r,echo=FALSE}
figures(1)
```

_Figure 1. Pallid sturgeon weight (y-axis) and length (x-axis) for 2003 to 2016 grouped by Missouri River
basin (Upper; pink open circles,lower; blue open circles)._  

```{r,echo=FALSE}
figures(2)# plot of mean weight at length for RPMA 2 and 4
```  

_Figure 2. Plot of weight (y-axis) and length (x-axis) on log scale for Pallid Sturgeon captured in Missouri River
basins (Upper; pink open circles, lower; blue open circles). 
Lines represent best fit linear models predicting weight from length for data in each panel and group._

### Patterns in condition

Figure 3 below illustrates boxplots of condition among years and basins.
Of interest to the skinny fish analysis are the years post 2011 that show a decline in condition.

```{r,echo=FALSE}
figures(3)# boxplot of Kn given mean weight at length for RPMA 2 and 4
``` 
_Figure 3. Plot of condition (y-axis) and day of year (x-axis) for Pallid Sturgeon captured in
the upper and lower basins._


### What are the predicted weights for a given length given fitted model
Evaluating plots of predicted weight for a given length by year and basin in Figure 4 below we 
generally see a similar view as Figure 3 above.





```{r,echo=FALSE}
figures(4)# mean weight for lengths
```  

_Figure 4. Plot of estimated weight (y-axis) and year (x-axis) for varying pallid sturgeon lengths (800, 1000, 1200, and 1400 mm).
Lower basin estimates were not included for fish 1400 mm because they were not observed.
Vertical lines denote 95% confidence intervals. The dotted horizonatal line represents the standard
weight used to calculate condition._ 


### Does mean kn correlate with predicted weights at varying sizes?

It should follow that condition (averaged by year and basin) and predicted weight for a 
given length by year and basin should exhibit similar patterns.  Figure 3 above ploted condition
by year and basin, which is similar to previous analyses presented to MRRIC and the ISAP and 
Figure 4 illustrated how the fitted model was used to estimate the weight for a given length
pallid sturgeon. Superimposing the average condition factors on the data from Figure 4 above 
for the lower and upper basin in Figures 5 and 6 below shows a bit of a discrepancy among some of
the patterns of average condition and predicted weight at length.  



```{r,echo=FALSE}
figures(5)
``` 

_Figure 5. _

```{r,echo=FALSE}
figures(6)
``` 

_Figure 6. _

```{r,echo=FALSE}
figures(7)
``` 

_Figure 7. Estimated condition values for upper basin pallid sturgeon of varying length
in 2015 and 2016.  The dotted lines illustrate relative frequency of observations._


